
\section{Setting}

Assume a state machine for each individual process. 

Each process state is a pair $(cs, ps)$ where $cs$~is a control state and $ps$
is a vector of parameters.  The first parameter of a component is always its
own identity.  Other parameters are references to other components.  For the
fixed process, all parameters are references to components.  Write $p.cs$ and
$p.\params$ for the control state and parameters of process state~$p$.  If $p$
is a component state, write $p.\id$ for its identity.
  
%%  We say that $s$ \emph{references} $id'$ if $id'$ is a parameter of~$s$
%% other than its own identity.

Write a system state as $s = (\fixed, \cpt)$ where $\fixed$ is the state of
the fixed process, and $\cpt$ is a set of states of components.  Write $\S$
for the set of all system states.  Write $s.\fixed$ and $s.\cpt$ for the fixed
processes or components, respectively.

If $s$ is a system state and $cpts$ is a set of component states with identities
disjoint from $s.\cpt$, then write $s \uplus cpts$ for $(s.\fixed, s.\cpt \union
cpts)$.  If $cpt$ is a component state, we abuse notation and write $cpt \in
s$ for $cpt \in s.\cpt$.

Assume different components have different identities for now. 

%% If $id$ is an identity, we say that $id$ is in $s$ if there is a
%% component~$cpt \in s$ such that $s.\id = id$.  We say that $s$ references $id$
%% if a component of~$s$ references~$id$. 

Assume that components are partitioned into \emph{active} and \emph{passive}
components.  E.g.~threads and nodes.  Each transition involves precisely one
active component or an active fixed process. 

\improve{check this in implementation.}
